Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-12-10 manuscript/usses_water_ms.Rmd
more tiny revisions and matching up with response letter

---
title: Ecosystem functional response across precipitation extremes in a sagebrush steppe
preprint: false
author: 
  - name: Andrew T. Tredennick
    affiliation: 1,*
    corresponding: true
    email: atredenn@gmail.com
  - name: Andrew R. Kleinhesselink
    affiliation: 1,2,*
  - name: J. Bret Taylor
    affiliation: 3
  - name: Peter B. Adler
    affiliation: 1
affiliation:
  - code: 1
    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
  - code: 3
    address: United States Department of Agriculture, Agricultural Research Service, U.S. Sheep Experiment Station, Dubois, Idaho 83423
abstract: |
  **Background.** Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common. Periods of extreme precipitation are major drivers of interannual variability in ecosystem functioning in water limited communities, but how ecosystems respond to these extremes over the long-term may shift with precipitation means and variances. Long-term changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.

@@

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}
\blfootnote{*Authors contributed equally to this work.}

@@

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemesia tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata}.
During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.

@@

\new{We then modeled these daily deviations ($\Delta x_\text{trt.}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions. 
Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
The day immediately following rainfall was also classified as rainy.
We fit the model using the ```lme4::lmer()``` function [@Bates2015] in R [@R2016], with random effects for plot group and date.
We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
This procedure allowed us to predict daily soil moisture conditions for all plots, even on those days when some of our direct treatment measurements were missing due to malfunction of the sensors or dataloggers. 
See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.

@@

For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS (Moderate Resolution Imaging Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer) bands for NDVI.
We regressed the greenness index against the dry biomass weight from the ten calibration plots to convert the greenness index to ANPP.
We fit regressions to a MODIS-based index and an AVHRR-based index for each year and retained the regression with the better fit based on $R^2$ values.
We then predicted ANPP using the best regression equation for each year (Appendix 1).
\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 2).}

@@

\noindent{}where $\Sigma$ is the variance-covariance matrix and $j(i)$ reads as "plot $j$ associated with observation $i$". 
The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-Cauchy distribution. 
\new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
A full description of our model and associated R [@R2016] code is in Appendix 2.

@@

\new{We conducted the above analysis for all species, and then conducted separate analyses for perennial species only and annual species only. Annual species have shorter life spans, so conducting separate analyses allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes. Given the dominance of perennial species in our system, shifts in the annual plant community could be masked in the analysis of the full community.}

@@

All R code and data necessary to reproduce our analysis has been archived on Figshare (\emph{link here after acceptance}) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
We also include annotated Stan code in our model description in Appendix 2.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3B).
The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that \new{broadly} overlap zero (Table 1).
Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig. A2-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the \new{slopes for the drought and} irrigation treatments are less than the slope for the control treatment.
This evidence comes from interpreting the posterior distributions of the slope offsets for the treatments.
\new{From these distributions,} we calculate a \new{91}% one-tailed probability that the estimate is less than zero for the irrigation treatment \new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment} (Fig. 3A, right panel).

\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. x). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. x)}.

Community composition was similar among treatments.
The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 4).
Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
\new{Splitting the analysis apart for perennials and annuals led to similar results.
The perennial community was stable over time and consistently similar among treatments (Fig. Ax-x).
There is some evidence that the annual community changed in response to treatment in two years (Fig. Ax-x), but these responses appear to come from very rare species since our analysis weights them equally (Appendix X).}

@@

First, it could be that our precipitation manipulations were not large enough to induce a response.
That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
We cannot definitively rule out this possibility, but we have reason to believe our treatments \emph{should} have been large enough.
Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig. A4-1).
The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 4).
Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3A).

Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
Failure to account for potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig. A2-2).
Unfortunately, we did not collect data on additional factors because our focus was on precipitation.

@@

Another strategy to deal with variable environmental conditions is avoidance, which would also result in a consistent ecosystem functional response between drought and control treatments.
For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
Furthermore, the dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
\new{The dominance of our site by the shrub \emph{A. tripartita} may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations} [e.g., @Huxman2004; @Knapp2015; @Wilcox2017].

\new{We did find that ANPP became more sensitive to drought over time (Fig. x).
Thus, the difference between ANPP, standardize by the difference in soil moisture, in drought and control plots grew over time.
The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
Ecosystem functional response may soon shift if sensitivity continues to increase.}

@@

Our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime, \new{at least over the five years of our experiment}. 
Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016]. 
For example, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply take more time than we observed here.}

@@

\begin{description}
\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
\item [Appendix 2.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
\item [Appendix 3.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
\item [Appendix 4.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
\end{description}

@@

```{r table, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
library(xtable)
effect_summary <- readRDS("../results/effect_summary.RDS")
colnames(effect_summary) <- c("Coefficient", "Treatment", "Posterior Mean", "Posterior Median",
                              "Lower 95\\% BCI", "Upper 95\\% BCI")
coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment."
print(xtable(effect_summary, caption = coef_caption, digits = 2),
      caption.placement="top",
      include.rownames = F,
      sanitize.colnames.function = identity,
      comment = FALSE,
      size = "normalsize")

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
\end{figure}

>>>>

---
title: Ecosystem functional response across precipitation extremes in a sagebrush steppe
preprint: false
author: 
  - name: Andrew T. Tredennick
    affiliation: 1
    corresponding: true
    email: atredenn@gmail.com
  - name: Andrew R. Kleinhesselink
    affiliation: 1,2
  - name: J. Bret Taylor
    affiliation: 3
  - name: Peter B. Adler
    affiliation: 1
affiliation:
  - code: 1
    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
  - code: 3
    address: United States Department of Agriculture, Agricultural Research Service, U.S. Sheep Experiment Station, Dubois, Idaho 83423
abstract: |
  **Background.** Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common. Periods of extreme precipitation are major drivers of interannual variability in ecosystem functioning in water limited communities, but how ecosystems respond to these extremes over the long-term may shift with precipitation means and variances. Long-term changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.

@@

<!---
\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}
\blfootnote{*Authors contributed equally to this work.}
--->

@@

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemesia tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because demographic data is difficult to collect and is already being collected in these plots for other ongoing studies.}

@@

\new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions. 
Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
The day immediately following rainfall was also classified as rainy.
We fit the model using the ```lme4::lmer()``` function [@Bates2015] in R [@R2016], with random effects for plot group and date.
We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
This procedure allowed us to predict daily soil moisture conditions for all plots, even on those days when some of our direct treatment measurements were missing due to malfunction of the sensors or dataloggers. 
See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.

@@

For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS (Moderate Resolution Imaging Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer) bands for NDVI.
We regressed the greenness index against the dry biomass weight from the ten calibration plots to convert the greenness index to ANPP.
We fit regressions to a MODIS-based index and an AVHRR-based index for each year and retained the regression with the better fit based on $R^2$ values.
We then predicted ANPP using the best regression equation for each year (Appendix 2).
\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 4).}

@@

\noindent{}where $\Sigma$ is the variance-covariance matrix and $j(i)$ reads as "plot $j$ associated with observation $i$". 
The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-Cauchy distribution. 
\new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
A full description of our model and associated R [@R2016] code is in Appendix 3.

@@

\new{We conducted the above analysis for all species, and then conducted a separate analyses for annual species only (Appendix 1). 
Annual species have shorter life spans, so conducting a separate analysis allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes. 
Given the dominance of perennial species in our system (Appendix 1), shifts in the annual plant community could be masked in the analysis of the full community.}

@@

All R code and data necessary to reproduce our analysis has been archived on Figshare (\emph{link here after acceptance}) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
We also include annotated Stan code in our model description in Appendix 3.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3B).
The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that \new{broadly} overlap zero (Table 1).
Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig. A3-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the \new{slopes for the drought and} irrigation treatments are less than the slope for the control treatment.
This evidence comes from interpreting the posterior distributions of the slope offsets for the treatments.
\new{From these distributions,} we calculate a \new{91}% one-tailed probability that the estimate is less than zero for the irrigation treatment \new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment} (Fig. 3A, right panel).

\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3D). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3D)}.

Community composition was similar among treatments.
The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 4).
Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
\new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
Splitting the analysis apart for annuals led to similar results.
There is some evidence that the annual community changed in response to treatment in two years (Fig. A1-6), but these responses appear to come from very rare species since our analysis weights them equally (Fig. A1-4).}

@@

First, it could be that our precipitation manipulations were not large enough to induce a response.
That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
We cannot definitively rule out this possibility, but we have reason to believe our treatments \emph{should} have been large enough.
Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig. A5-1).
The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 5).
Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3A).

Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
\new{For example,} @LaPierre2016 \new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
Moreover, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
Failure to account for potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig. A3-2).
Unfortunately, we did not collect data on additional factors because our focus was on precipitation.

@@

Another strategy to deal with variable environmental conditions is avoidance, which would also result in a consistent ecosystem functional response between drought and control treatments.
For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
Furthermore, the dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
\new{The dominance of our site by the shrub \emph{A. tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations} [e.g., @Huxman2004; @Knapp2015; @Wilcox2017].

\new{We did find that ANPP became more sensitive to drought over time (Fig. 3D).
Thus, the difference between ANPP, standardized by the difference in soil moisture, in drought and control plots grew over time.
The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
Ecosystem functional response may soon shift if sensitivity continues to increase.}

@@

Our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime, \new{at least over the five years of our experiment}. 
Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016], \new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.} 
Likewise, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply takes more time than we observed here.}

@@

\begin{description}
\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Table A1-2, Table A1-3, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, Fig. A1-5, and Fig. A1-6.}
\item [Appendix 2.] Additional methods and information on estimating aboveground net primary productivity, Table A2-1, and Table A2-2.
\item [Appendix 3.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
\item [Appendix 4.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
\item [Appendix 5.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
\end{description}

@@

```{r table, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
library(xtable)
effect_summary <- readRDS("../results/effect_summary.RDS")
colnames(effect_summary) <- c("Coefficient", "Treatment", "Posterior Mean", "Posterior Median",
                              "Lower 95\\% BCI", "Upper 95\\% BCI")
coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment ($\\beta$ coefficients in equation 1). The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment. Posterior distributions of the offsets are in Figure 3A."
print(xtable(effect_summary, caption = coef_caption, digits = 2),
      caption.placement="top",
      include.rownames = F,
      sanitize.colnames.function = identity,
      comment = FALSE,
      size = "normalsize")

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment. \new{Each point represents a plot and the axes represent species composition.}}
\end{figure}
